\chapter{Introduction}

The objective of this thesis is to present some applications of machine learning and deep learning in systems research.
 Specifically, this thesis is organized as follows.

Chapter \ref{chap:water} presents an application of Machine Learning in edge computing systems research.
 The Edge Computing application here deals with improving the performance of water softeners and improving water utility efficiency.
 The approach taken is to engineer appropriate features so that water hardness classification and forecasting is reduced to a linear classification task that would be computationally light to be programmed on edge computing systems (wireless routers etc.).

Chapter \ref{chap:spectrum} presents an application of Machine Learning in wireless systems research.
 The wireless systems research application here deals with estimating wireless spectrum usage from sparse measurements across a geographic area (a small city in this case).
 The approach taken is to use collaborative filtering and estimate the missing spectrum measurements across time at a specific location.
 In other words, the approach attempts to characterize the time dynamics of wireless spectrum usage.

Chapter \ref{chap:sound} presents an application of Deep Learning in mobile systems research.
 The mobile systems research application here deals with preserving user-privacy in developing voice-enabled mobile applications.
 The approach taken is to investigate lightweight audio obfuscation techniques that maintain good keyphrase recognition performance.

Chapter \ref{chap:conclusion} presents conclusions and directions for future research.
